Multidimensional information

### 2022-09-28 notebooks/Multidimensional_information.qmd
Minor corrections to multidimensional infromation notebook

@ackerman_creating_1994's Equations 1 through 7 and 13 through 16 for
bidimensional latent spaces apply directly to the 3-or-more dimensional case,
regardless of the latent covariance structure, as this is not involved at all in
those expressions. The only caveat is that he uses the normal parameterization,
and thus for the M2PL as we have been formulating it the 1.7 factor in Equation
1 does not apply (and subsequently, the 1.7\^2 = 2.89 factor in Equations 7 and
14).

@@

I will first try with a correlated, bidimensional latent space as an example to
explore the solution. In this space, the information of simple-structure items
in the direction parallel to the axes of its tapped trait should be equal to the
information of the unidimensional case. In the direction of the other trait
axis, the information should be more positive the more positive the correlation
is (constrained by the information of the former axis as maximum value).

I first define the latent space, by considering a bivariate standardize normal
distribution with correlated dimensions:

@@

These are the cosines, the components of the vector $\beta$. The covariance
error matrix, Equation 13 in @ackerman_creating_1994, is given by the inverse of
the information matrix. The latter can be expressed as
$\mathbf{A}^T\mathbf{A}PQ$, being $\mathbf{A}$ the matrix of all item
parameters; the covariance matrix can be subsequently expressed as
$(\mathbf{A}^T\mathbf{A})^{-1}PQ^{-1}$. In our case, for a single item,
$\mathbf{A} = a^T$.

>>>>

@ackerman_creating_1994's Equations 1 through 7 and 13 through 16 for
bidimensional latent spaces apply directly to the 3-or-more dimensional case,
regardless of the latent covariance structure, as this is not involved at all in
those expressions. The only caveat is that he uses the normal parameterization,
and thus for the M2PL as we have been formulating it the 1.7 factor in Equation
1 does not apply (and subsequently, the 1.7^2^ = 2.89 factor in Equations 7 and
14).

@@

I will first try with a correlated, bidimensional latent space as an example to
explore the solution. In this space, the information of simple-structure items
in the direction parallel to the axes of its tapped trait should be equal to the
information of the unidimensional case. In the direction of the other trait
axis, the information should be higher the more positive the correlation is
(constrained by the information of the former axis as maximum value).

I first define the latent space, by considering a bivariate standardized normal
distribution with correlated dimensions:

@@

These are the cosines, the components of the vector $\beta$. The error
covariance matrix, Equation 13 in @ackerman_creating_1994, is given by the
inverse of the information matrix. The latter can be expressed as
$\mathbf{A}^T\mathbf{A}PQ$, being $\mathbf{A}$ the matrix of all item
parameters; the error covariance matrix can be subsequently expressed as
$(\mathbf{A}^T\mathbf{A})^{-1}PQ^{-1}$. In our case, for a single item,
$\mathbf{A} = a^T$.
